Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-11-11 Cost_manuscript.Rmd
rewording of Results sections up until modeling

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey, Gary Kane, & Joseph T. McGuire"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: #default
    reference_docx: Style_reference_draft.docx
  html_document:
    df_print: paged
csl: apa.csl
#subtitle: 'Abbreviated Title: Heterogeneity in the functional topography of mPFC'
#bibliography: mPFC_paper.bib
---

@@

  In summary, the free parameters from the subjective time model captured three features. First, how fast or slow a given demand felt ($S_{demand}$). Second, whether and how fast this time sensitivity was recalibrated based on the experience of intermixed demands ($\alpha_w$). Finally, whether this time sensitivity was used retrospectively (i.e. high $\alpha$), or prospectively (i.e. low $\alpha$).

@@

  We performed a mixed effects logistic regression to address the first hypothesis. Since participants in the effort groups were forced to travel when they failed to perform above threshold (see Methods), we distinguish between acceptances (entering the handling period) and completions (successfully obtaining a reward). Larger reward amounts significantly increased acceptance probabilities ($\beta$ = `r historyCoeffs_btw["Offer", "Estimate"]`, SE = `r historyCoeffs_btw["Offer", "Std. Error"]`, *p* `r report_p(historyCoeffs_btw["Offer", "Pr(>|z|)"])`), whereas longer delays decreased them ($\beta$ = `r historyCoeffs_btw["Handling", "Estimate"]`, SE = `r historyCoeffs_btw["Handling", "Std. Error"]`, *p* `r report_p(historyCoeffs_btw["Handling", "Pr(>|z|)"])`), in line with predictions. Two additional regressors showed that having missed out on consecutive rewards decreased the probability of acceptance ($\beta$ = `r historyCoeffs_btw["recentTravels", "Estimate"]`, SE = `r historyCoeffs_btw["recentTravels", "Std. Error"]`, *p* `r report_p(historyCoeffs_btw["recentOffers", "Pr(>|z|)"])`), and that participants became more selective when recent offer history was richer ($\beta$ = `r historyCoeffs_btw["recentOffers", "Estimate"]`, SE = `r historyCoeffs_btw["recentOffers", "Std. Error"]`, *p* `r report_p(historyCoeffs_btw["recentOffers", "Pr(>|z|)"])`). The reduction in acceptances from recent misses was driven by the wait group (removing this group from the model yielded a *p* > 0.1), likely because this group accepted the least amount of offers. Comparing likelihood fits among lengths of offer history (e.g. sum over 2 or 3 past trials) suggested that participants were influenced by the `r historyIndex_btw + 1` most recent rewards (AIC = `r round(AIC(historySize_btw[[historyIndex_btw]]), digits = 2)`), which matched the offer presentation rotation (see Methods).

@@

  The next set of analyses examined the hypothesis that choice differences among costs would be characterized by more acceptances by the effort groups. Figure 3 (left) shows that the cognitive effort group consistently accepted more offers, which resulted in lower earnings (middle). One-way ANOVAs showed that differences among groups were significant both for overall acceptance rates (F(`r propAov[[1]]$Df[1]`, `r propAov[[1]]$Df[2]`) = `r round(propAov[[1]][1, 4], digits = 2)`, *p* `r report_p(propAov[[1]][1, 5])`), and total earnings (F(`r earnAov[[1]]$Df[1]`, `r earnAov[[1]]$Df[2]`) = `r round(earnAov[[1]][1, 4], digits = 2)`, *p* `r report_p(earnAov[[1]][1, 5])`). Post-hoc permutations (5000 iterations) comparing mean acceptance rates between every pair of costs confirmed that acceptance rates of the cognitive effort group (mean = `r filter(overallSumms, Cost == "Cognitive")$mProp`, SD = `r filter(overallSumms, Cost == "Cognitive")$sdProp`) were higher than those of the physical (mean = `r filter(overallSumms, Cost == "Physical")$mProp`, SD = `r filter(overallSumms, Cost == "Physical")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Physical < Cognitive")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Physical < Cognitive")$cohenD`) and wait groups (mean = `r filter(overallSumms, Cost == "Wait")$mProp`, SD = `r filter(overallSumms, Cost == "Wait")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Cognitive > Wait")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Cognitive > Wait")$cohenD`). It also showed that those faced with the easy task accepted more than those in the wait group, although responses from the easy group were more variable (mean Easy = `r filter(overallSumms, Cost == "Easy")$mProp`, SD = `r filter(overallSumms, Cost == "Easy")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Wait < Easy")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Wait < Easy")$cohenD`) (all other *p* > 0.05). The proportion of forced travels for cognitive effort was `r round(propFails[2, 2], digits = 2)` across participants (and 0 for either physical or easy conditions), making performance an unlikely explanation for the observed differences in choice. The higher acceptance rates produced by the cognitive effort group resulted in significantly lower earnings (mean = `r filter(overallSumms, Cost == "Cognitive")$mEarn`, SD = `r filter(overallSumms, Cost == "Cognitive")$sdEarn`) than both physical (mean = `r filter(overallSumms, Cost == "Physical")$mEarn`, SD = `r filter(overallSumms, Cost == "Physical")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Physical > Cognitive")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Physical > Cognitive")$cohenD`) and wait groups (mean = `r filter(overallSumms, Cost == "Wait")$mEarn`, SD = `r filter(overallSumms, Cost == "Wait")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Cognitive < Wait")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Cognitive < Wait")$cohenD`). In addition, we found that the wait group earned more than the physical (*p* `r report_p(filter(earnTests, contrast == "Physical < Wait")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Physical < Wait")$cohenD`) and easy groups (mean Easy = `r filter(overallSumms, Cost == "Easy")$mEarn`, SD = `r filter(overallSumms, Cost == "Easy")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Wait > Easy")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Wait > Easy")$cohenD`).

  We next examined the optimality of these decisions. The foraging task was configured so that there was a single reward-maximizing strategy per block type (see Operationalization of Cost in Methods), and participants were warned that accepting everything would not maximize their gains. Accordingly, participants who over- or under-accepted earned the least (Figure 3, right; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, Beta = `r round(summary(lmEarn)$coefficients[3,1], digits = 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, R-squared = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). Two-sided, one-sample t-tests against the optimal proportion of acceptances (i.e. $\mu$ = 0 or 1) for each combination of cost type, handling time, and offer showed that participants significantly deviated from optimality in `r round(optimalProp)`% of combinations (*p* < 0.05, FDR corrected), with the percentage of deviations being lowest for the wait group (Wait = `r filter(optimalProp_cost, Cost == "Wait")[2]`%, physical = `r filter(optimalProp_cost, Cost == "Physical")[2]`%, Cognitive = `r filter(optimalProp_cost, Cost == "Cognitive")[2]`%, Easy = `r filter(optimalProp_cost, Cost == "Easy")[2]`%). Together, these findings show that even though participants did not always perform optimally, they adequately integrated the statistics of the environment as they tracked their recent experience. Further, they suggest that cognitive effort can boost the value of an offer, as reflected by higher acceptances at the expense of earnings.

@@

```{r mixed effects plots (Fig. 4), out.width = "80%", fig.align = "center", fig.cap = "Comparing acceptance rates among costs for each reward and handling combination. A:  B:  Each entry shows how much more likely was the reference group to accept an offer than the comparison group. C:"}
#hrPlot_btw / (ABIC_plot_btw | ME_matrix_btw) # doesn't work with corrplot.. oh well
include_graphics("./Images/FIG4.png")
```

  We next probed whether these results held uniformly across all parameter combinations. Figure 4A shows the mean proportion of acceptances (± SEM) per combination of handling time, reward, and cost, with optimal acceptance rates depicted by the gray circles. Qualitatively, the figure confirms two important predictions. First, participants adapted to the richness of the each timing block, showing a pattern of acceptances that resembled the optimal strategy regardless of the cost they faced. Second, differences in cost were consistent across blocks, and acceptances were highest for the cognitive effort group. The optimality of the strategies produced by each cost depended on the richness of the block. This can help explain why the wait group earned the most despite having similar acceptance rates to the physical group, as the former was closer to optimality in the richest environment. We tested these observations with a mixed-effects logistic regression, computing the probability of accepting an offer as a function of handling time, reward amount, and demand type (see Methods for details). As before, the model showed significant main effects of handling time (global intercept = `r round(fixef(mixLogis_main$Wait)["(Intercept)"], digits = 2)`, $\beta$ = `r round(fixef(mixLogis_main$Wait)["Handling"], digits = 2)`, SE = `r round(summary(mixLogis_main$Wait)[[10]]["Handling", "Std. Error"], digits = 2)`, *p* `r report_p(summary(mixLogis_main$Wait)[[10]]["Handling", "Pr(>|z|)"])`) and offer amount ($\beta$ = `r round(fixef(mixLogis_main$Wait)["Offer"], digits = 2)`, SE = `r round(summary(mixLogis_main$Wait)[[10]]["Offer", "Std. Error"], digits = 2)`, *p* `r report_p(summary(mixLogis_main$Wait)[[10]]["Offer", "Pr(>|z|)"])`). Comparisons among all conditions are shown on Figure 4B, which portrays the coefficients (color scale), and significance that resulted from switching the reference cost condition ("X" marks non-significant comparisons). The matrix of coefficients conveys that the cognitive effort group was significantly more likely to accept any offer than the physical ($\beta$ = `r round(mixSummary_btw$Betas["Physical", "Cognitive"], digits = 2)`, *p* `r report_p(mixSummary_btw$Pvals["Physical", "Cognitive"])`) and wait groups ($\beta$ = `r round(mixSummary_btw$Betas["Wait", "Cognitive"], digits = 2)`, *p* `r report_p(mixSummary_btw$Pvals["Wait", "Cognitive"])`), but not the easy group.

  These observations could not be better explained by simpler or more complex regression models. Figure 4C shows AIC and BIC values for models of increasing complexity, starting with an intercept-only configuration (see Methods for details). Here we focus on AIC, as both metrics yielded comparable results. As predicted, the a priori model with all main effects performed better than those relying on a single parameter (a priori model $R^2$ = `r round(mixLogis_compare$AllMainRsq, digits = 2)`), as well as the model with main effects for handling time and reward amount (AIC~a-priori~ = `r filter(ME_evals$abicLogis_compare, Model == "AllMain")$AIC`, AIC~handling/reward~ = `r filter(ME_evals$abicLogis_compare, Model == "HR")$AIC`, $\chi^2$(`r ME_evals$HRvMain[[7]][2]`) = `r round(ME_evals$HRvMain$Chisq[2], digits = 2)`, *p* `r report_p(ME_evals$HRvMain[[8]][2])`). However, against our prediction, the a priori model was outperformed by models that added an interaction between handling and reward (AIC~HR_interaction~ = `r filter(ME_evals$abicLogis_compare, Model == "HR_interaction")$AIC`, $\chi^2$(`r ME_evals$MainvHRint[[7]][2]`) = `r round(ME_evals$MainvHRint$Chisq[2], digits = 2)`, *p* `r report_p(ME_evals$MainvHRint[[8]][2])`) as well as one that considered all possible interactions (AIC~all_interactions~ = `r filter(ME_evals$abicLogis_compare, Model == "Interaction")$AIC`, $\chi^2$(`r ME_evals$MainvAllint[[7]][2]`) = `r round(ME_evals$MainvAllint$Chisq[2], digits = 2)`, *p* `r report_p(ME_evals$MainvAllint[[8]][2])`). These improvements were driven by the fact that differences in offer acceptances among costs were never present for 20 cents (all *p* values for interactions containing offer amount < 0.001). As the highest reward should always be accepted in prey foraging environments, we conclude that the most parsimonious explanation of choice behavior comes from the a priori model containing all main effects. In combination with the pattern observed in Figure 4A, these results confirm that handling and reward amounts were similarly integrated across groups, but that the absolute rate of acceptances was affected by the cost faced (and was highest for those confronted with cognitive demands).

@@

  Next, we tested the hypothesis that foragers would display consistent and stable choices. Participants were deemed consistent if they did not quit after engaging in a trial. To examine this, we plotted survival curves of trial quits across all participants (Figure 5, top-left), with black crosses signaling either completed trials (at 2 s, 10 s, and 14 s) or forced travels. These curves show that most decisions to quit happened within the first second into the handling time, and participants rarely quit during the handling time. Physical and easy groups show slightly lagged responses because the experiment allowed a one-second grace period for participants to begin gripping (marked by the dashed line). This resulted in some individuals choosing to wait for that second to indicate an offer rejection (the censor at 4s in the physical group was a completion during a 2 s handling trial that was incorrectly recorded as 4 s by the experimental code). Plotting the quitting time distributions across blocks for the first handling second (ECDFs on Figure 5, top-right) shows that participants in the wait and cognitive groups responded faster as the experiment progressed. These response time observations suggest clarity in participant choices

@@

```{r Prepost mixed effects plot (Fig. 6), fig.align = "center", fig.cap = "Distilling acceptance behavior before and after the break. Left: Proportion of acceptances per handling, reward, and half, showing that effortful demands were increasingly discounted over time. Right: Mixed-effects coefficients denoting the comparison among costs acceptances for each half separately. Even though acceptances steadily declined in the effort groups, the relative preferences were preserved."}
include_graphics("./Images/FIG6.png")

@@

  If the cost or value of a demand were intrinsic, the relative preferences displayed so far should be robust to the ability to obtain the same rewards through different demands. We investigated this possibility by modifying the foraging task, such that every participant experienced every type of demand in blocks that interleaved wait trials with either physical or cognitive effort (fixed at 10 s handling and 6 s travel times, to allow for comparisons between experiments). We refer to observations from each block's wait trials by their effort pair (i.e. "Wait-C" for cognitive, and "Wait-P" for physical). We hypothesized that 1) reward decision patterns would match those of the previous experiment, and 2) that interleaving effort and delay would produce persistent preferences for wait trials.

@@

  Similar to the between-subject results, a survival analysis displayed that participants made most of their choices quickly within the offer window (Figure 8A). The median percentage of quits during the handling time was `r round(median(propQuits_wth$propAccept), digits = 2)`% (SD = `r round(sd(propQuits_wth$propAccept), digits = 2)`), with only `r sum(propQuits_wth$propAccept > 20)` participants quitting over 20% of trials during the handling time. Moreover, the cumulative quitting time distributions show that participant responses during the offer window became faster over time (Figure 8B). Once in the handling time, the overall proportion of forced travels in the effortful trials was relatively low for both cognitive (median = `r round(median(filter(propFails_wth, Cost == "Cognitive")$propFails), digits = 2)`%, SD = `r round(sd(filter(propFails_wth, Cost == "Cognitive")$propFails), digits = 2)`) and physical (median = `r round(median(filter(propFails_wth, Cost == "Physical")$propFails), digits = 2)`%, SD = `r round(sd(filter(propFails_wth, Cost == "Physical")$propFails), digits = 2)`) conditions. Similar to the between-subjects experiment, these results suggest that decision makers were consistent in their adopted strategies, and were able to perform the tasks well.

>>>>

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey, Gary Kane, & Joseph T. McGuire"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
output:
  word_document: #default
    reference_docx: Style_reference_draft.docx
  pdf_document:
    highlight: haddock
    keep_tex: no
  html_document:
    df_print: paged
csl: apa.csl
#subtitle: 'Abbreviated Title: Heterogeneity in the functional topography of mPFC'
#bibliography: mPFC_paper.bib
---

@@

  In summary, the free parameters from the subjective time model captured three features. First, how fast or slow a given demand felt ($S_{demand}$). Second, how early this time sensitivity was recalibrated and integrated into decisions based on the experience of intermixed demands ($\alpha_w$). Finally, whether this time sensitivity was used retrospectively (i.e. high $\alpha$), or prospectively (i.e. low $\alpha$).

@@

  We performed a mixed effects logistic regression to address the first hypothesis. Since participants in the effort groups were forced to travel when they failed to perform above threshold (see Methods), we distinguish between acceptances (entering the handling period) and completions (successfully obtaining a reward). In line with our predictions, larger reward amounts significantly increased acceptance probabilities ($\beta$ = `r historyCoeffs_btw["Offer", "Estimate"]`, SE = `r historyCoeffs_btw["Offer", "Std. Error"]`, *p* `r report_p(historyCoeffs_btw["Offer", "Pr(>|z|)"])`), whereas longer delays decreased them ($\beta$ = `r historyCoeffs_btw["Handling", "Estimate"]`, SE = `r historyCoeffs_btw["Handling", "Std. Error"]`, *p* `r report_p(historyCoeffs_btw["Handling", "Pr(>|z|)"])`). Two additional regressors showed that having missed out on consecutive rewards decreased the probability of acceptance ($\beta$ = `r historyCoeffs_btw["recentTravels", "Estimate"]`, SE = `r historyCoeffs_btw["recentTravels", "Std. Error"]`, *p* `r report_p(historyCoeffs_btw["recentOffers", "Pr(>|z|)"])`), and that participants became more selective when recent offer history was richer ($\beta$ = `r historyCoeffs_btw["recentOffers", "Estimate"]`, SE = `r historyCoeffs_btw["recentOffers", "Std. Error"]`, *p* `r report_p(historyCoeffs_btw["recentOffers", "Pr(>|z|)"])`). The reduction in acceptances from recent misses was driven by the wait group (removing this group from the model yielded a *p* > 0.1), likely because this group accepted the least amount of offers. Comparing likelihood fits among lengths of offer history (e.g. sum over 2 or 3 past trials) suggested that participants were influenced by the `r historyIndex_btw + 1` most recent rewards (AIC = `r round(AIC(historySize_btw[[historyIndex_btw]]), digits = 2)`), which matched the offer presentation rotation (see Methods).

@@

  The next set of analyses examined the hypothesis that effort group would uniformly accept more trials. Figure 3 (left) shows that the cognitive effort group consistently accepted more offers, which resulted in lower earnings (middle). One-way ANOVAs showed that differences among groups were significant both for overall acceptance rates (F(`r propAov[[1]]$Df[1]`, `r propAov[[1]]$Df[2]`) = `r round(propAov[[1]][1, 4], digits = 2)`, *p* `r report_p(propAov[[1]][1, 5])`), and total earnings (F(`r earnAov[[1]]$Df[1]`, `r earnAov[[1]]$Df[2]`) = `r round(earnAov[[1]][1, 4], digits = 2)`, *p* `r report_p(earnAov[[1]][1, 5])`). Post-hoc permutations (5000 iterations) comparing mean acceptance rates between every pair of costs confirmed that acceptance rates of the cognitive effort group (mean = `r filter(overallSumms, Cost == "Cognitive")$mProp`, SD = `r filter(overallSumms, Cost == "Cognitive")$sdProp`) were higher than those of the physical (mean = `r filter(overallSumms, Cost == "Physical")$mProp`, SD = `r filter(overallSumms, Cost == "Physical")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Physical < Cognitive")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Physical < Cognitive")$cohenD`) and wait groups (mean = `r filter(overallSumms, Cost == "Wait")$mProp`, SD = `r filter(overallSumms, Cost == "Wait")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Cognitive > Wait")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Cognitive > Wait")$cohenD`). It also showed that those faced with the easy task accepted more than those in the wait group, although responses from the easy group were more variable (mean Easy = `r filter(overallSumms, Cost == "Easy")$mProp`, SD = `r filter(overallSumms, Cost == "Easy")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Wait < Easy")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Wait < Easy")$cohenD`) (all other *p* > 0.05). The higher acceptance rates produced by the cognitive effort group resulted in significantly lower earnings (mean = `r filter(overallSumms, Cost == "Cognitive")$mEarn`, SD = `r filter(overallSumms, Cost == "Cognitive")$sdEarn`) than both physical (mean = `r filter(overallSumms, Cost == "Physical")$mEarn`, SD = `r filter(overallSumms, Cost == "Physical")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Physical > Cognitive")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Physical > Cognitive")$cohenD`) and wait groups (mean = `r filter(overallSumms, Cost == "Wait")$mEarn`, SD = `r filter(overallSumms, Cost == "Wait")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Cognitive < Wait")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Cognitive < Wait")$cohenD`). In addition, we found that the wait group earned more than the physical (*p* `r report_p(filter(earnTests, contrast == "Physical < Wait")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Physical < Wait")$cohenD`) and easy groups (mean Easy = `r filter(overallSumms, Cost == "Easy")$mEarn`, SD = `r filter(overallSumms, Cost == "Easy")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Wait > Easy")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Wait > Easy")$cohenD`). The proportion of forced travels for cognitive effort was `r round(propFails[2, 2], digits = 2)` across participants, and 0 for either physical or easy conditions, making performance an unlikely explanation for the observed differences in choice.

  We next examined the optimality of these decisions. The foraging task was configured so that there was a single reward-maximizing strategy per block type (see Operationalization of Cost in Methods), and participants were warned that accepting everything would not maximize their gains. Accordingly, participants who over- or under-accepted earned the least (Figure 3, right; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, Beta = `r round(summary(lmEarn)$coefficients[3,1], digits = 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, $R^2$ = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). Two-sided, one-sample t-tests against the optimal proportion of acceptances (i.e. $\mu$ = 0 or 1) for each combination of cost type, handling time, and offer showed that participants significantly deviated from optimality in `r round(optimalProp)`% of combinations (*p* < 0.05, FDR corrected), with the percentage of deviations being lowest for the wait group (Wait = `r filter(optimalProp_cost, Cost == "Wait")[2]`%, physical = `r filter(optimalProp_cost, Cost == "Physical")[2]`%, Cognitive = `r filter(optimalProp_cost, Cost == "Cognitive")[2]`%, Easy = `r filter(optimalProp_cost, Cost == "Easy")[2]`%). Together, these findings show that even though participants did not always perform optimally, they adequately integrated the statistics of the environment as they tracked their recent experience. Further, they suggest that cognitive effort can boost the value of an offer, as reflected by higher acceptances at the expense of earnings.

@@

```{r mixed effects plots (Fig. 4), out.width = "80%", fig.align = "center", fig.cap = "Comparing acceptance rates among costs for each reward and handling combination. A: Proportion accepted for every combination of handling time, reward, and demand. Grey dots indicate the reward-maximizing acceptance for any handling/reward combination. B: Matrix portraying the coefficients (color scale) and significance that resulted from switching the reference cost condition ("X" marks non-significant comparisons). Each entry shows how much more likely a group was to accept compared to the reference condition (row). C: AIC and BIC values for a forward mixed-effects model comparison, ranging from a base intercept-only to a full-interaction model."}
#hrPlot_btw / (ABIC_plot_btw | ME_matrix_btw) # doesn't work with corrplot.. oh well
include_graphics("./Images/FIG4.png")
```

  We next probed whether these results held uniformly across all parameter combinations. Figure 4A shows the mean proportion of acceptances (± SEM) per combination of handling time, reward, and cost, with optimal acceptance rates depicted by the gray circles. Qualitatively, the figure confirms two important predictions. First, participants adapted to the richness of the each timing block, gravitating towards optimality regardless of the cost they faced. Second, differences in cost were consistent across blocks, and acceptances were uniformly highest for the cognitive effort group. We tested these observations with a mixed-effects logistic regression, computing the probability of accepting an offer as a function of handling time, reward amount, and demand type (see Methods for details). As before, the model showed significant main effects of handling time (global intercept = `r round(fixef(mixLogis_main$Wait)["(Intercept)"], digits = 2)`, $\beta$ = `r round(fixef(mixLogis_main$Wait)["Handling"], digits = 2)`, SE = `r round(summary(mixLogis_main$Wait)[[10]]["Handling", "Std. Error"], digits = 2)`, *p* `r report_p(summary(mixLogis_main$Wait)[[10]]["Handling", "Pr(>|z|)"])`) and offer amount ($\beta$ = `r round(fixef(mixLogis_main$Wait)["Offer"], digits = 2)`, SE = `r round(summary(mixLogis_main$Wait)[[10]]["Offer", "Std. Error"], digits = 2)`, *p* `r report_p(summary(mixLogis_main$Wait)[[10]]["Offer", "Pr(>|z|)"])`). Comparisons among all conditions are shown on Figure 4B. The matrix of coefficients conveys that the cognitive effort group was significantly more likely to accept any offer than the physical ($\beta$ = `r round(mixSummary_btw$Betas["Physical", "Cognitive"], digits = 2)`, *p* `r report_p(mixSummary_btw$Pvals["Physical", "Cognitive"])`) and wait groups ($\beta$ = `r round(mixSummary_btw$Betas["Wait", "Cognitive"], digits = 2)`, *p* `r report_p(mixSummary_btw$Pvals["Wait", "Cognitive"])`), but not the easy group.

  These observations could not be better explained by simpler or more complex regression models. Figure 4C shows AIC and BIC values for models of increasing complexity, starting with an intercept-only configuration (see Methods for details). Here we focus on AIC, as both metrics yielded comparable results. As predicted, the a priori model with all main effects performed better than those relying on a single parameter (a priori model $R^2$ = `r round(mixLogis_compare$AllMainRsq, digits = 2)`), as well as the model with main effects for handling time and reward amount (AIC~a-priori~ = `r filter(ME_evals$abicLogis_compare, Model == "AllMain")$AIC`, AIC~handling/reward~ = `r filter(ME_evals$abicLogis_compare, Model == "HR")$AIC`, $\chi^2$(`r ME_evals$HRvMain[[7]][2]`) = `r round(ME_evals$HRvMain$Chisq[2], digits = 2)`, *p* `r report_p(ME_evals$HRvMain[[8]][2])`). However, against our expectations, the a priori model was outperformed by models that added an interaction between handling and reward (AIC~HR_interaction~ = `r filter(ME_evals$abicLogis_compare, Model == "HR_interaction")$AIC`, $\chi^2$(`r ME_evals$MainvHRint[[7]][2]`) = `r round(ME_evals$MainvHRint$Chisq[2], digits = 2)`, *p* `r report_p(ME_evals$MainvHRint[[8]][2])`) as well as one that considered all possible interactions (AIC~all_interactions~ = `r filter(ME_evals$abicLogis_compare, Model == "Interaction")$AIC`, $\chi^2$(`r ME_evals$MainvAllint[[7]][2]`) = `r round(ME_evals$MainvAllint$Chisq[2], digits = 2)`, *p* `r report_p(ME_evals$MainvAllint[[8]][2])`). These improvements were driven by the fact that differences in offer acceptances among costs were never present for 20 cents (all *p* values for interactions containing offer amount < 0.001). As the highest reward should always be accepted in prey foraging environments, we conclude that the most parsimonious explanation of choice behavior came from the a priori model containing all main effects. These results further support the hypotheses that handling and reward amounts were similarly integrated across groups, and that the absolute rate of acceptances was affected by the demand faced. However, they offer partial support for the hypothesis that effort would induce higher acceptance rates, which was only true for those confronted with cognitive effort.

@@

  Next, we tested the hypothesis that foragers would display consistent and stable choices. Participants were deemed consistent if they did not quit after engaging in a trial. To examine this, we plotted survival curves of trial quits across all participants (Figure 5, top-left), with black crosses signaling either completed trials (at 2 s, 10 s, and 14 s) or forced travels. These curves show that most decisions to quit happened within the first second into the handling time, and participants rarely quit during the handling time. Physical and easy groups showed slightly lagged responses because the experiment allowed a one-second grace period for participants to begin gripping (marked by the dashed line). This resulted in some individuals opting to wait for that second to indicate an offer rejection (the censor at 4s in the physical group was a completion during a 2 s handling trial that was incorrectly recorded as 4 s by the experimental code). Plotting the quitting time distributions across blocks for the first second of the handling time (ECDFs on Figure 5, top-right) showed that participants in the wait and cognitive groups responded faster as the experiment progressed. These response time observations suggested clarity in participant choices

@@

```{r Prepost mixed effects plot (Fig. 6), fig.align = "center", fig.cap = "Distilling acceptance behavior before and after the break. Left: Proportion of acceptances per handling, reward, and half, showing that effortful demands became increasingly aversive over time. Right: Mixed-effects coefficients denoting the comparison among costs acceptances for each half separately (mirrored along the diagonal). Even though acceptances steadily decreased in the effort groups, the relative preferences were preserved."}
include_graphics("./Images/FIG6.png")

@@

  If the cost or value of a demand were intrinsic, the relative preferences displayed so far should have been robust to the ability to obtain the same rewards through different demands. We investigated this possibility by modifying the foraging task, such that every participant experienced every type of demand in blocks that interleaved wait trials with either physical or cognitive effort (fixed at 10 s handling and 6 s travel times, to allow for comparisons between experiments). We refer to observations from each block's wait trials by their effort pair (i.e. "Wait-C" for cognitive, and "Wait-P" for physical). We hypothesized that 1) reward decision patterns would match those of the previous experiment; and 2) that interleaving effort and delay would produce persistent preferences for wait trials.

@@

  Similar to the between-subject results, a survival analysis displayed that participants made most of their choices quickly within the offer window (Figure 8A). The median percentage of quits during the handling time was `r round(median(propQuits_wth$propAccept), digits = 2)`% (SD = `r round(sd(propQuits_wth$propAccept), digits = 2)`), with only `r sum(propQuits_wth$propAccept > 20)` participants quitting over 20% of trials during the handling time. Moreover, the cumulative quitting time distributions show that participant responses during the offer window became faster over time (Figure 8B). Once in the handling time, the overall percentage (**shift to percentage in exp 1**) of forced travels in the effortful trials was low for both cognitive (median = `r round(median(filter(propFails_wth, Cost == "Cognitive")$propFails), digits = 2)`%, SD = `r round(sd(filter(propFails_wth, Cost == "Cognitive")$propFails), digits = 2)`) and physical (median = `r round(median(filter(propFails_wth, Cost == "Physical")$propFails), digits = 2)`%, SD = `r round(sd(filter(propFails_wth, Cost == "Physical")$propFails), digits = 2)`) conditions. Similar to the between-subjects experiment, these results suggest that decision makers were consistent in their adopted strategies, and were able to perform the tasks well.
